---
title: "Homework1"
author: "Yi Chen"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Use the built-in dataset trees to do the following:
### a) Obtain the density plot of the Height variable.
```{r message=FALSE, warning=FALSE}
data(trees)
attach(trees)
plot(density(trees$Height), main="Density plot of the hight")
```

### b) Obtain a scatterplot matrix of all 3 variables.
```{r}
pairs(trees[, 1:3])

```

### c) Obtain the mean vector, the variance-covariance matrix ($\textit{S}_n$), and the correlation matrix of all 3 variables.
```{r}
colMeans(trees)
cov(trees)
cor(trees)
```
## 2. Do exercise 3 on p. 38.

```{r}
x1 <- c(9,2,6,5,8)
x2 <- c(12,8,6,4,10)
x3 <- c(3,4,0,2,1)
X <- cbind(x1,x2,x3)
```
The array of measure mean is
```{r}
apply(X,2,mean)

```
The variance-covariance matrix is
```{r}
cov(X)
```
Notice: the covariance in this function is based on n - 1 degrees of freedom instead of n. 


The correlation matrix is
```{r}
cor(X)
```



