---
title: 'R-Lab: Fixed income Securities'
author: "Yi Chen"
date: "1/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R-Lab: Fixed income Securities
```{r}
# C = cupon payment(semiannual)
# T = time to maturity(in years)
# r = vector of yields to maturity (semiannual rates)
# F = face value / par value
# P = bond prices corresponding to all values of yield to maturity in the input vector r

bondprice = function(C,T,r,F){
  P = C / r + (F - C / r) * (1 + r)^(-2 * T)
}

```
Suppose we need to compute the yield to maturity for a 30-year par 1,000 bond with coupon payments of 40 selling at 1,200.

We could calculate the bond prices for a grid of interest rates using the previous function.

The price-yield function is then interpolated to find the yield to maturity when the price is 1,200.
```{r}
P = 1200 # current price of the bond 
C = 40 # cupon payment
T = 30 # time to maturity
F = 1000 # par value of the bond

r = seq(0.02,0.05,length=300) # grid of interest rates
value = bondprice(C,T,r,F)
yield2m = spline(value,r,xout = P) # spline interpolation
yield2m

```
```{r}
plot(r, value, xlab = "yield to maturity", ylab = "bond price",
type = "l", main = "F = 1000, C = 40, T = 30", lwd = 2)
abline(h = 1200)
abline(v = yield2m)
```
As an alternative, R has a built-in nonlinear root finder
uniroot(),which can be used to solve the yield to maturity directly.
uniroot()solves for the root where the equation equals zero.
In the above example, it is equivalent to solve for the root of r when the difference of the bond price function and the current price is 0.
```{r}
uniroot(function(r) bondprice(C,T,r,F)-P, c(.02,.05))
```

Note that in the first argument, the function has to solely depend onr.  And we have to specify the range where the root lies inside in the second argument.
