---
title: "Homework12"
author: "Yi Chen"
date: "4/19/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## homework 12

#### F test

$$F_{obs}=\frac{\Delta R^2/(PA-PC)}{(1-R^2_A)/(N-PA)}$$


Using the F test to compare the extended tree model over the additive tree model.
According to the output of sport data, $N=170$ (number of pair of proximities), $PC=37$, $PA=10$, $R^2_A=0.9306$, and $R^2_B=0.8172$.
Consequently,
$$F_{obs}=\frac{(0.9306-0.8172)/(37-9)}{(1-0.9306)/(170-37)}=7.76$$
```{r}
1- pf(q = 7.76,df1 = 37-9,df2 = 170-37)
```

The P-value is almost 0, which indicates that the model has a significant imporve over the addtive tree.



#### Interpretation

The additive tree summary 10 subset in clustering solution. Based on the characteristic of each subset, we can interprete it as:

C: table sports

D: sports with slow movemen

E: high body contact and fighting

H: sports related to quick and consistant jumping. 

I: water sports

N: slide on a surface

O: related to hunting

U: outdoor sports

X: ball sports

By looking at the tree, we find that some of the sports have multiple marked features, like surfing (I/N), canoeing (I/U), camping (O/U), boxing (E/H) and archery (DOU). Those sports share two or more common points with other sports. For example, canoeing is a water sports (like sports in I) and also an outdoor sport (like sports in U).