---
title: "Homework3"
author: "Yi Chen"
date: "2/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

$$ \mathbf{x} \sim N(\mathbf{\mu},\mathbf{\Sigma})$$
$$\mathbf{\mu} = [0,0,0]' , \mathbf{\Sigma}=\begin{bmatrix} \sigma^2,0,0 \\ 0,\sigma^2,0 \\ 0,0,\sigma^2 \end{bmatrix}$$
The linear combination of normal distribution is still normal distribution. 
$$E(w)=E(\frac{x_1+x_2+x_3}{3})=\frac{1}{3}(\mu_1+\mu_2+\mu_3)=0$$
$$Var(w)=Var(\frac{x_1+x_2+x_3}{3})=\frac{1}{9}(\sigma_1^2+\sigma_2^2+\sigma_3^2)=\frac{1}{3}\sigma^2$$
In summary

$$w \sim N(0,\frac{1}{3}\sigma^2)$$

## Problem 2

### a)
```{r}
data <- airquality
qqnorm(data$Solar.R)
qqline(data$Solar.R)
```

Firstly, the QQ plot indicates that the data is not fully match with the normal distribution. 

### b)
```{r}
shapiro.test(data$Solar.R)
```

Secondly, Shapiro-Wilks test has very small p-value, which also reject normality.

## Problem 3

### a)

```{r}
x.bar <- c(2.9,0.9,2.9)
mu.0 <- c(3,1,4)
sigma <- matrix(c(6,1,-2,1,13,4,-2,4,4), 3, 3)
n <- 14
p <- 3
z.obs <- n*t(x.bar - mu.0)%*%solve(sigma)%*%(x.bar - mu.0)
1-pchisq(z.obs,df=p)
```

Based on the level of significance $\alpha=0.01$, the hyphothesis cannot be rejected. 

### b)

```{r}
z.obs1 <- n*(x.bar[1]-mu.0[1])^2/sigma[1,1]
z.obs2 <- n*(x.bar[2]-mu.0[2])^2/sigma[2,2]
z.obs3 <- n*(x.bar[3]-mu.0[3])^2/sigma[3,3]
1-pchisq(z.obs1,df=1)
1-pchisq(z.obs2,df=1)
1-pchisq(z.obs3,df=1)
```

Based on the level of significance $\alpha=0.01$, none of the three individuals hyphothesis' can be rejected. 



